How has the Brazilian Amazon been constructed as a problem:Presidential speeches and transnational politics since 1985

### 2022-09-20 Draft_September_2022/Full_draft_20220901.rmd
#tabmodelin

---
title: How has the Brazilian Amazon been constructed as a problem:Presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: 20 September 2022"
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\text{Word Count:}#1} 
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: Presidential speeches influence the ways we think about and act towards the environment transnationally, but they are understudied. We propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in discourses across time and space. Using supervised machine learning, we classify statements about the Amazon in 6130 presidential speeches since 1985. We find that national and international events drive the frequency at which the topic of the Amazon is mentioned in presidential speeches inside and outside Brazil. While constructing the Amazon as a problem of economic integration dominates discourses until the mid-2000s, environmental conservation and social development constructions increase in relevance and temporarily surpass economic integration from 2010 to 2015. In turn, constructing the Amazon as an issue of sovereignty steadily increases since 2010. Lastly, the farthest away presidents are from the Amazon itself, the more likely they are to construct the Amazon as an issue of environmental conservation.
---

```{=latex}
\keywords{ discourse analysis, transnational governance, environmental policy, Brazilian Amazon, supervised learning, deforestation}
```
```{=latex}
\wordcount{ 8470 (including abstract, text, references, and and footnotes)}
```
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(tidyr)
library(scales)
library(RColorBrewer)
library(stargazer)
library(kableExtra)
library(tibble)
library(readxl)
library(ggplot2)
library(gridExtra)
Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

@@

We conceptualize four problem-constructions: sovereignty, economic integration, social development, and conservation. Figure 4, below, illustrates the share of each pure-type in time. Pure problem-constructions dominate, with an average of 56% of all Amazonian statements across time.We identify a discursive shift in problem-constructions from developing the Amazon economically, to conserving it and providing services to its citizens. Pure economic integration constructions dominated constructions from 1985 to the late 2000s. This is especially pertinent during the Cardoso administration. Brazilian presidents after Cardoso moved from employing one problem-constructions consistently to a more diverse strategy. During Lula's terms in office, constructing the Amazon purely as an issue of economic integration decreased, while constructing the Amazon issues of environmental conservation and social development increased. Conservation and social development constructions surpassed economic integration around 2010 and continued to dominate as Amazonian constructions during Rousseff's administration. Still, environmental conservation and social development problem-constructions were not as dominant in terms of frequency or periods in time in discourses as economic integration was until the early 2000s. Finally, we observe a steady increase in sovereignty constructions in the late 2000s, reaching the highest frequencies with the presidencies of Bolsonaro.As we conceptualize and operationalize sovereignty as containing internal and external perceived threats to the Amazon, we interpret this increase as not only attacks on international interference in the Amazon but also on indigenous and traditional populations.

@@

```{r Figure 6, fig.cap="Logistic Regression predicted values", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}
ama_model <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") # merge
ama_model <- dplyr::left_join(ama_model, AAI, by = "year") # merge
ama_model <- filter(ama_model, location_cat != "Non Identified",
                    mx_cat != "Other Mixed-types") %>%
  mutate(con_vs_all = ifelse(mx_cat == "Pure Environmental Conservation", 1, 0),
         EI_vs_all = ifelse(mx_cat == "Pure Economic Integration", 1, 0),
         SD_vs_all = ifelse(mx_cat == "Pure Social Development", 1, 0),
         sov_vs_all = ifelse(mx_cat == "Pure National Sovereignty", 1, 0))

@@

# Plot model
ama_model$"Environmental Conservation" = predict(model_logit_con, ama_model, type = "response")
ama_model$"Economic Integration" = predict(model_logit_ei, ama_model, type = "response")
ama_model$"Social Development" = predict(model_logit_sd, ama_model, type = "response")
ama_model$"National Sovereignty" = predict(model_logit_sov, ama_model, type = "response")
plot_loc <- ama_model %>% gather(key=p_c, value = pred, 22:24)
plot_loc$p_c <-factor(plot_loc$p_c, levels = c("Economic Integration",
                                               "Environmental Conservation",
                                               "Social Development"))
all_loc <- ggplot(plot_loc, aes(x = km_to_manaus, y=pred, color=p_c)) +
  geom_jitter(alpha=1, size=1) +
  geom_smooth(size = .5, se=FALSE) +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Distance from the Amazon (km)",
       y = "",
       title = "Predicted probability of each problem-construction as a function of\ndistance from the Amazon",
       subtitle="Sovereignty was removed because of statistical insignificance (table 1).", 
       caption = "Curves in the plots are estimated using loess method.") +
    theme_am() +
  theme(legend.position = "none")+
  facet_wrap(~p_c, ncol=3)

@@

Table 1, below, shows that for each 1000Km increase, we find an increase of 52.8% in the predicted probability of constructing the Amazon as an issue of environmental preservation. That is, the farthest away the president is from the Amazon region, the more likely the president is to construct it as an issue of environmental conservation. Second,the opposite is true for both economic integration (and social development): for an increase of 1000Km, we find a decrease of 48.5% (47.4%), respectively. When speaking within the Amazon, presidents focus on social development and economic integration. This entails that conservation might be a desirable construction when speaking internationally about the Amazon, but not at the local level, where presidents highlight economic and social development. This multi-level game implies that presidents talk to the economic and social needs of the people domestically while boasting about environmentalism internationally.

@@

logit2prob(coef(model_logit_con2))
logit2prob(coef(model_logit_ei2))
logit2prob(coef(model_logit_sd2))

@@

>>>>

---
title: How has the Brazilian Amazon been constructed as a problem:Presidential speeches and transnational politics since 1985
author:
- Livio Silva-Muller
- Henrique Sposito^[Both co-authors contributed equally to the article. Names are ordered alphabetically. Livio is a PhD Candidate in Anthropology and Sociology at the Geneva Graduate Institute and affiliated to the Albert Hirschman Centre on Democracy. Henrique is a PhD Candidate in International Relations and Political Sciences at the Geneva Graduate Institute and affiliated to the Centre for International Environmental Studies.]
date: 20 September 2022
output:
  pdf_document: default
  latex_engine: xelatex
  fig_caption: yes
  html_document:
    df_print: paged
fontsize: 12pt
mainfont: Times New Roman
header-includes:
   - \usepackage{floatrow}
   - \floatsetup[figure]{capposition=top}
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \usepackage[T1]{fontenc}
   - \usepackage[utf8]{inputenc}
   - |
    \providecommand{\keywords}[1]{\textbf{Keywords:} #1}
    \providecommand{\wordcount}[1]{\text{Word Count:}#1} 
always_allow_html: TRUE
bibliography: references.bib
csl: chicago.csl
abstract: Presidential speeches influence the ways we think about and act towards the environment transnationally, but they are understudied. We propose a framework to investigate how the Brazilian Amazon has been constructed as a problem in discourses across time and space. Using supervised machine learning, we classify statements about the Amazon in 6130 presidential speeches since 1985. We find that national and international events drive the frequency at which the topic of the Amazon is mentioned in presidential speeches inside and outside Brazil. While constructing the Amazon as a problem of economic integration dominates discourses until the mid-2000s, environmental conservation and social development constructions increase in relevance and temporarily surpass economic integration from 2010 to 2015. In turn, constructing the Amazon as an issue of sovereignty steadily increases since 2010. Lastly, the farthest away presidents are from the Amazon itself, the more likely they are to construct the Amazon as an issue of environmental conservation.
---

```{=latex}
\keywords{ discourse analysis, transnational governance, environmental policy, Brazilian Amazon, supervised learning, deforestation}
```
```{=latex}
\wordcount{ 8470 (including abstract, text, references, and and footnotes)}
```
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(tidyr)
library(scales)
library(RColorBrewer)
library(stargazer)
library(kableExtra)
library(tibble)
library(readxl)
library(ggplot2)
library(gridExtra)
library(sjPlot)

Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

@@

We conceptualize four problem-constructions: sovereignty, economic integration, social development, and conservation. Figure 4, below, illustrates the share of each pure-type in time. Pure problem-constructions dominate, with an average of 56% of all Amazonian statements across time.We identify a discursive shift in problem-constructions from developing the Amazon economically, to conserving it and providing services to its citizens. Pure economic integration constructions dominated constructions from 1985 to the late 2000s. This is especially pertinent during the Cardoso administration. Brazilian presidents after Cardoso moved from employing one problem-constructions consistently to a more diverse strategy. During Lula's terms in office, constructing the Amazon purely as an issue of economic integration decreased, while constructing the Amazon issues of environmental conservation and social development increased. Conservation and social development constructions surpassed economic integration around 2010 and continued to dominate as Amazonian constructions during Rousseff's administration. Still, environmental conservation and social development problem-constructions were not as dominant in terms of frequency or periods in time in discourses as economic integration was until the early 2000s. Finally, we observe a steady increase in sovereignty constructions in the late 2000s, reaching the highest frequencies with the presidencies of Bolsonaro. As we conceptualize and operationalize sovereignty as containing internal and external perceived threats to the Amazon, we interpret this increase as not only attacks on international interference in the Amazon but also on indigenous and traditional populations.

@@

```{r Figure 6, fig.cap="Logistic Regression predicted values", echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, include=TRUE}

ama_model <- dplyr::left_join(ama_mx, amazon_def_year, by = "year") # merge
ama_model <- dplyr::left_join(ama_model, AAI, by = "year") # merge
ama_model <- filter(ama_model, location_cat != "Non Identified",
                    mx_cat != "Other Mixed-types") %>%
  mutate(con_vs_all = ifelse(mx_cat == "Pure Environmental Conservation", 1, 0),
         EI_vs_all = ifelse(mx_cat == "Pure Economic Integration", 1, 0),
         SD_vs_all = ifelse(mx_cat == "Pure Social Development", 1, 0),
         sov_vs_all = ifelse(mx_cat == "Pure National Sovereignty", 1, 0))

@@

# Plot model
ama_model$"Environmental Conservation" = predict(model_logit_con, ama_model, type = "response")
ama_model$"Economic Integration" = predict(model_logit_ei, ama_model, type = "response")
ama_model$"Social Development" = predict(model_logit_sd, ama_model, type = "response")
ama_model$"National Sovereignty" = predict(model_logit_sov, ama_model, type = "response")
plot_loc <- ama_model %>% gather(key=p_c, value = pred, 22:25)

plot_loc$p_c <-factor(plot_loc$p_c, levels = c("Economic Integration",
                                               "Environmental Conservation",
                                               "Social Development",
                                               "National Sovereignty"))
all_loc <- ggplot(plot_loc, aes(x = km_to_manaus, y=pred, color=p_c)) +
  geom_jitter(alpha=1, size=1) +
  geom_smooth(size = .5, se=FALSE) +
  scale_y_continuous(labels = percent_format()) +
  labs(x = "Distance from the Amazon (km)",
       y = "",
       title = "Predicted probability of each problem-construction as a function of\ndistance from the Amazon",
       caption = "Curves in the plots are estimated using loess method.") +
    theme_am() +
  theme(legend.position = "none")+
  facet_wrap(~p_c, ncol=2)

@@

We convert log odds from Table 1, below, into probabilities [^2]. It shows that for each 1000Km increase, we find an increase of 52.8% in the predicted probability of constructing the Amazon as an issue of environmental preservation. That is, the farthest away the president is from the Amazon region, the more likely the president is to construct it as an issue of environmental conservation. Second,the opposite is true for both economic integration (and social development): for an increase of 1000Km, we find a decrease of 48.5% (47.4%), respectively. When speaking within the Amazon, presidents focus on social development and economic integration. This entails that conservation might be a desirable construction when speaking internationally about the Amazon, but not at the local level, where presidents highlight economic and social development. This multi-level game implies that presidents talk to the economic and social needs of the people domestically while boasting about environmentalism internationally. Finally, national sovereignty problem constructions do not appear to change with distance.

[^2 See table 3 in appendix for all odds ratios and confidence intervals.]

@@

logit2prob(coef(model_logit_con2))
logit2prob(coef(model_logit_ei2))
logit2prob(coef(model_logit_sd2))
logit2prob(coef(model_logit_sov2))

@@

```{r Table 3, fig.cap="Table 3: Odds Ratios and Confidence Intervals", echo=FALSE, message=TRUE, warning=TRUE, paged.print=TRUE, include= TRUE}

tab_model(model_logit_con2,model_logit_ei2,model_logit_sd2,model_logit_sov2)
